---
title: "Getting Start with Assisted CIDER (asCIDER)"
description: "Hu et al. An interpretable meta-clustering framework for single-cell RNA-Seq data integration and evaluation"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting Start with Assisted CIDER (asCIDER)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This vignette performs asCIDER, a meta-clustering method, on a cross-species pancreas dataset. AsCIDER is aimed to achieve the clustering task in data confounded by unwanted variables. In this scenario, the unwanted variable is species effects.

asCIDER is short for assisted CIDER, assisted by the prior batch-specific annotations or clustering results.

# Set up

In addition to **CIDER**, we will load the following packages:

```{r setup, message=FALSE}
library(CIDER)
library(Seurat)
library(pheatmap)
library(ggplot2)
library(cowplot)
library(viridis)
```

# Pancreas data

The example data can be downloaded from https://figshare.com/s/d5474749ca8c711cc205. 

This data$^1$ contain cells from human (8241 cells) and mouse (1886 cells).

```{r} 
load("../data/pancreas_counts.RData") # count matrix
load("../data/pancreas_meta.RData") # meta data/cell information
seu <- CreateSeuratObject(counts = pancreas_counts, meta.data = pancreas_meta)
table(seu$Batch)
```

# Exam if data are confounded

Prior to use CIDER (or other integration methods for clustering) it is important to exam if clustering will be confounded by the cross-species factors.


Her we first perform the conventional Seurat$^2$ clustering pipeline.

```{r seurat-pipeline}
seu <- NormalizeData(seu, verbose = FALSE)
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
seu <- ScaleData(seu, verbose = FALSE)
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunTSNE(seu, reduction = "pca", dims = 1:12)
```

## Confounded dimension reduction 

The PCA and *t*-SNE plots both showed that the data are confounded by species. 

The `scatterPlot` function is used to generate dimension reduction figures. It takes input of a Seurat object (`seu` here), the name of reduction (`pca` and `tsne` here), the variable deciding dot colours (`Batch` here) and the title of plots. See more information by `?scatterPlot`

```{r scatterPlot-confounded, fig.width=7, fig.height=3}
p1 <- scatterPlot(seu, "pca",colour.by = "Batch", title = "PCA") 
p2 <- scatterPlot(seu, "tsne",colour.by = "Batch", title = "t-SNE") 
plot_grid(p1, p2)
```

## Confounded clustering results

```{r, fig.width=5, fig.height=3}
seu <- FindNeighbors(seu, dims = 1:12)
seu <- FindClusters(seu)

scatterPlot(seu, "tsne",colour.by = "seurat_clusters", title = "t-SNE") 
```

```{r, fig.width=5, fig.height=3}
res <- data.frame(table(seu$seurat_clusters, seu$Batch))
ggplot(res, aes(fill=Var2, y=Freq, x=Var1)) + 
    geom_bar(position="stack", stat="identity") + xlab("CIDER_cluster") + ylab("Proportions")
```


# asCIDER

## Prepare initial clusters

asCIDER uses existing within-batch clustering results. Here we concatenate the batch ID and the within-batch cluster ID to obtain cluster-specific groups (i.e. initial clusters).

```{r initial_cluster}
seu$initial_cluster <- paste(seu$Group, seu$Batch, sep = "_")
table(seu$initial_cluster)
```

## Calculate of IDER similarity matrix

The function `getIDEr` calculate the IDER-based distance matrix. 

By default, it will use the column called "initial_cluster" as initial clusters, and "Batch" as batch. If you are using columns other than this two, please revise these two parameters.

For this step, you can choose to use parallel computation by setting `use.parallel = TRUE` or not (default `use.parallel = FALSE`). The default number of cores used for parallel computation is `detectCores(logical = FALSE) - 1`.

```{r run-getIDER-noparallel}
ider <- getIDEr(seu, 
                group.by.var = "initial_cluster",
                batch.by.var = "Batch",
                downsampling.size = 35, 
                use.parallel = FALSE, verbose = FALSE)
```

## Visualise the similarity matrix

```{r heatmap, fig.height=7, fig.width=7}
groups <- c("alpha","beta","delta", "gamma","ductal","endothelial", "activated_stellate", "quiescent_stellate", "macrophage")
idx1 <- paste0(groups, "_human")
idx2 <- paste0(groups, "_mouse")

pheatmap::pheatmap(
  ider[[1]][idx1, idx2],
  color = inferno(10),
  border_color = NA,
  display_numbers = TRUE,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  width = 7,
  height = 5,
  cellwidth = 22,
  cellheight = 22
)
```


## Perform final Clustering

Next we put both Seurat object and the similarity matrix list into `finalClustering`. You can either cut trees by height (set `cutree.by = 'h`) or by k (set `cutree.by = 'k`). The default is cutting by the height of 0.45.

The final clustering results are stored in Colume `cider_final_cluster` of Seurat object metadata and can be extracted using `seu$cider_final_cluster`.

```{r final-clustering}
seu <- finalClustering(seu, ider, cutree.h = 0.45)
head(seu@meta.data)
```

# Visualise clustering results

```{r summary-tsne-plots, fig.width=8, fig.height=3}
plot_list <- list()
plot_list[[1]] <- scatterPlot(seu, "tsne", colour.by = "CIDER_cluster", title = "asCIDER clustering results") 
plot_list[[2]] <- scatterPlot(seu, "tsne", colour.by = "Group", title = "Ground truth of cell populations") 
plot_grid(plotlist = plot_list, ncol = 2)
```

```{r, fig.height=3, fig.width=5}
res <- data.frame(table(seu$CIDER_cluster, seu$Batch))
ggplot(res, aes(fill=Var2, y=Freq, x=Var1)) + 
    geom_bar(position="stack", stat="identity") + xlab("CIDER_cluster") + ylab("Proportions")
```

# Technical

```{r sessionInfo}
sessionInfo()
```


# References

1. Baron, M. et al. A Single-Cell Transcriptomic Map of the Human and Mouse Pancreas Reveals Inter- and Intra-cell Population Structure. Cell Syst 3, 346–360.e4 (2016).
2. Satija R, et al. Spatial reconstruction of single-cell gene expression data. Nature Biotechnology 33, 495-502 (2015). 
