---
title: "Getting Start with De Novo CIDER (dnCIDER)"
output: 
  rmarkdown::html_vignette:
    toc: TRUE
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Getting Start with De Novo CIDER (dnCIDER)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

This vignette performs dnCIDER on a cross-species pancreas dataset.


# Set up

In addition to **CIDER**, we will load the following packages:

```{r setup}
library(CIDER)
library(Seurat)
library(parallel)
library(cowplot)
```

# Load pancreas data

The example data can be downloaded from https://figshare.com/s/d5474749ca8c711cc205. 

Pancreatic cell data$^1$ contain cells from human (8241 cells) and mouse (1886 cells).

```{r} 
load("../data/pancreas_counts.RData") # count matrix
load("../data/pancreas_meta.RData") # meta data/cell information
seu <- CreateSeuratObject(counts = pancreas_counts, meta.data = pancreas_meta)
table(seu$Batch)
```

# Perform dnCIDER (high-level)

DnCIDER contains three steps 

```{r}
seu <- initialClustering(seu, additional.vars.to.regress = "Sample", dims = 1:15)
ider <- getIDEr(seu, downsampling.size = 35, use.parallel = FALSE, verbose = FALSE)
seu <- finalClustering(seu, ider, cutree.h = 0.35) # final clustering
```

# Visualise clustering results

We use the Seurat pipeline to perform normalisation (`NormalizeData`), preprocessing (`FindVariableFeatures` and `ScaleData`) and dimension reduction (`RunPCA` and `RunTSNE`).

```{r seurat-pipeline}
seu <- NormalizeData(seu, verbose = FALSE)
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
seu <- ScaleData(seu, verbose = FALSE)
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunTSNE(seu, reduction = "pca", dims = 1:12)
```

We can see 

```{r tsne-plot-CIDER-results, fig.height=3, fig.width=4}
scatterPlot(seu, "tsne", colour.by = "CIDER_cluster", title = "asCIDER clustering results") 
```

By comparing the dnCIDER results to the cell annotation from the publication$^1$, we observe that dnCIDER correctly identify the majority of populations across two species.

```{r tsne-plot-ground-truth, fig.height=3, fig.width=4}
scatterPlot(seu, "tsne", colour.by = "Group", title = "Ground truth of cell populations") 
```

# Technical

```{r sessionInfo}
sessionInfo()
```


# References

1. Baron, M. et al. A Single-Cell Transcriptomic Map of the Human and Mouse Pancreas Reveals Inter- and Intra-cell Population Structure. Cell Syst 3, 346–360.e4 (2016).
2. Satija R, et al. Spatial reconstruction of single-cell gene expression data. Nature Biotechnology 33, 495-502 (2015). 


