---
title: "Detailed walk-through of de novo CIDER (dnCIDER) on pancreas data"
output: 
  rmarkdown::html_vignette:
    toc: TRUE
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Detailed walk-through of de novo CIDER (dnCIDER) on pancreas data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction

This vignette performs dnCIDER on a cross-species pancreas dataset. It is aimed to show the underneath structure of dnCIDER compared to the other high level [vignette](docs/articles/dnCIDER_highlevel.html).

# Set up

In addition to **CIDER**, we will load the following packages:

```{r setup}
library(CIDER)
library(Seurat)
library(parallel)
library(cowplot)
```

# Load example data

The example data can be downloaded from https://figshare.com/s/d5474749ca8c711cc205. 

Pancreatic cell data$^1$ contain cells from human (8241 cells) and mouse (1886 cells).

```{r} 
load("../data/pancreas_counts.RData") # count matrix
load("../data/pancreas_meta.RData") # meta data/cell information
seu <- CreateSeuratObject(counts = pancreas_counts, meta.data = pancreas_meta)
table(seu$Batch)
```

# Perform initial clustering

```{r}
seu_list <- Seurat::SplitObject(seu, split.by = "Batch")
seu_list <- mclapply(seu_list, function(x) {
  x <- NormalizeData(x, normalization.method = "LogNormalize", 
                     scale.factor = 10000, verbose = FALSE)
  x <- FindVariableFeatures(x, selection.method = "vst", 
                            nfeatures = 2000, verbose = FALSE)
  x <- ScaleData(x, verbose = FALSE, vars.to.regress = "Sample")
  x <- RunPCA(x, features = VariableFeatures(object = x), verbose = FALSE)  
  x <- FindNeighbors(x, dims = 1:15, verbose = FALSE)
  x <- FindClusters(x, resolution = 0.6, verbose = FALSE)
  return(x)
})
```

```{r}
knitr::kable(table(seu_list[[2]]$Group, seu_list[[2]]$seurat_clusters))
```


```{r}
seu_list <- mclapply(seu_list, RunTSNE, dims = 1:15)
p1 <- scatterPlot(seu_list[[1]], "tsne", colour.by = "seurat_clusters")
p2 <- scatterPlot(seu_list[[2]], "tsne", colour.by = "seurat_clusters")
cowplot::plot_grid(p1,p2)
```

```{r}
dist_coef <- getDistMat(seu_list, downsampling.size = 50)
```


```{r}
par(mfrow = c(length(seu_list),1))
for(i in which(sapply(dist_coef, function(x) return(!is.null(x))))){
  tmp <- dist_coef[[i]] + t(dist_coef[[i]])
  diag(tmp) <- 1
  pheatmap::pheatmap(tmp, display_numbers = TRUE)
}
```

```{r}
for(seu_itor in 1:2){
  tmp <- dist_coef[[seu_itor]] + t(dist_coef[[seu_itor]])
  diag(tmp) <- 1
  tmp <- 1 - tmp
  hc <- hclust(as.dist(tmp), method = "average")
  hres <- cutree(hc, h = 0.4)
  df_hres <- data.frame(hres)
  df_hres$hres <- paste0(df_hres$hres, "_", unique(seu_list[[seu_itor]]$Batch))
  seu_list[[seu_itor]]$inicluster_tmp <- paste0(seu_list[[seu_itor]]$seurat_clusters, "_", seu_list[[seu_itor]]$Batch)
  seu_list[[seu_itor]]$inicluster <- df_hres$hres[match(seu_list[[seu_itor]]$inicluster_tmp,rownames(df_hres))]
}
```


```{r}
# plot(as.dendrogram(hc), horiz = T)
```

```{r, fig.height=3, fig.width=9}
p1 <- scatterPlot(seu_list[[1]], "tsne", "inicluster")
p2 <- scatterPlot(seu_list[[2]], "tsne", "inicluster")
plot_grid(p1,p2)
```

```{r}
scatterPlot(seu_list[[2]], "tsne", "Group")
```

# Calculate of IDER similarity matrix

```{r run-getIDER-parallel}
res <- unlist(lapply(seu_list, function(x) return(x$inicluster)))
res_names <- unlist(lapply(seu_list, function(x) return(colnames(x))))
seu$initial_cluster <- res[match(colnames(seu), res_names)]

ider <- getIDEr(seu, 
                group.by.var = "initial_cluster",
                batch.by.var = "Batch",
                downsampling.size = 35, 
                use.parallel = FALSE, verbose = FALSE)
```

```{r plot-network, fig.height=9, fig.width=9}
net <- plotNetwork(seu, ider, colour.by = "Group" , vertex.size = 0.6, weight.factor = 5)

```


```{r, fig.height=4}
hc <- hclust(as.dist(1-(ider[[1]] + t(ider[[1]])))/2)
plot(as.dendrogram(hc), horiz = TRUE)
```

# Perform final Clustering

```{r perform-final-clustering}
seu <- finalClustering(seu, ider, cutree.h = 0.35) # final clustering
```

```{r seurat-pipeline-for-visualisation}
seu <- NormalizeData(seu, verbose = FALSE)
seu <- FindVariableFeatures(seu, selection.method = "vst", 
                            nfeatures = 2000, verbose = FALSE)
seu <- ScaleData(seu, verbose = FALSE)
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunTSNE(seu, reduction = "pca", dims = 1:12)
```

```{r tsne-plot-CIDER-results, fig.height=3, fig.width=7}
plot_list <- list()
plot_list[[1]] <- scatterPlot(seu, "tsne", colour.by = "CIDER_cluster", title = "asCIDER clustering results") 
plot_list[[2]] <- scatterPlot(seu, "tsne", colour.by = "Group", title = "Ground truth of cell populations") 
plot_grid(plotlist = plot_list, ncol = 2)
```

# Technical

```{r sessionInfo}
sessionInfo()
```

# References

1. Baron, M. et al. A Single-Cell Transcriptomic Map of the Human and Mouse Pancreas Reveals Inter- and Intra-cell Population Structure. Cell Syst 3, 346–360.e4 (2016).
2. Satija R, et al. Spatial reconstruction of single-cell gene expression data. Nature Biotechnology 33, 495-502 (2015). 


