---
title: "Run meta-clustering (asCIDER) on the pancreas dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pancreas}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CIDER)
library(Seurat)
library(pheatmap)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(viridis)

qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
```


# Load data

Pancreatic cell data$^1$ contain cells from human (n = 8241) and mouse (1886).

```{r}
data("pancreas_counts")
data("pancreas_meta")
seu <- CreateSeuratObject(counts = pancreas_counts, meta.data = pancreas_meta)
table(seu$Batch)
```

# Prepare initial clusters

asCIDER uses existing within-batch clustering results. Here we concatenate the batch ID and the within-batch cluster ID to obtain cluster-specific groups (i.e. initial clusters).

```{r initial_cluster}
seu$initial_cluster <- paste(seu$Group, seu$Batch, sep = "_")
table(seu$initial_cluster)
```

# Calculation of IDER similarity matrix

The function `getIDEr` calculate the IDER-based similarity matrix. 

By default, it will use the column called "initial_cluster" as initial clusters, and "Batch" as batch. If you are using columns other than this two, please revise these two parameters.

The available DE methods are "trend" (default) for limma-trend and "voom" for limma-voom. limma-trend usually runs faster than voom and they tend to give similar outputs.

For this step, you can choose to use parallel computation by setting `use.parallel = TRUE` or not (default `use.parallel = FALSE`). The default number of cores used for parallel computation is `detectCores(logical = FALSE) - 1`.

```{r run-getIDER-parallel}
sim <- getIDEr(seu, 
               group.by.var = "initial_cluster",
               batch.by.var = "Batch",
               de.method = "trend",
               downsampling.size = 35)
```

Visualise the distance matrix with heatmap.

```{r heatmap, fig.height=7, fig.width=5}
groups <- c("alpha","beta","delta", "gamma","ductal","endothelial", "activated_stellate", "quiescent_stellate", "macrophage")
idx1 <- paste0(groups, "_human")
idx2 <- paste0(groups, "_mouse")
tmp <- sim[[1]] + t(sim[[1]])
diag(tmp) <- 1
pheatmap::pheatmap(
  tmp[idx1, idx2],
  color = inferno(10),
  border_color = NA,
  display_numbers = TRUE,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  width = 7,
  height = 5,
  cellwidth = 25,
  cellheight = 25
)
```


# Final Clustering

Next we put both Seurat object and the similarity matrix list into `finalClustering`. You can either cut trees by height (set `cutree.by = 'h`) or by k (set `cutree.by = 'k`). The default is cutting by the height of 0.45.

The final clustering results are stored in Colume `cider_final_cluster` of Seurat object metadata and can be extracted using `seu$cider_final_cluster`.

```{r final-clustering}
seu <- finalClustering(seu, sim, cutree.h = 0.45)
head(seu@meta.data)
```

# Visualisation

```{r seurat-pipeline}
seu <- NormalizeData(seu, verbose = FALSE)
seu <- FindVariableFeatures(seu, selection.method = "vst", nfeatures = 2000, verbose = FALSE)
seu <- ScaleData(seu, verbose = FALSE)
seu <- RunPCA(seu, npcs = 20, verbose = FALSE)
seu <- RunTSNE(seu, reduction = "pca", dims = 1:15)
```

```{r summary-tsne-plots, fig.width=6, fig.height=15}
df_plot <- data.frame(tSNE_1 = seu@reductions$tsne@cell.embeddings[,1],
                      tSNE_2 = seu@reductions$tsne@cell.embeddings[,2],
                      Batch = as.factor(seu$Batch),
                      Ground_truth = as.factor(seu$Group),
                      asCIDER_cluster = as.factor(seu$cider_final_cluster)
                      )

scatterPlot <- function(df, var, colvec) { # function for tSNE plot
  ggplot(df_plot, aes_string(x = "tSNE_1", y = "tSNE_2", col = var)) + 
    geom_point(alpha = 0.6, size=0.5)  +  
    theme_classic() + theme(legend.position = "right") +  
    theme(axis.ticks.x = element_blank(),axis.ticks.y = element_blank(),axis.text = element_blank()) +
    scale_color_manual(values = colvec)
}
  
plot_list <- list()
plot_list[[1]] <- scatterPlot(df = df_plot, var = "Batch", colvec = col_vector[1:10]) 
plot_list[[2]] <- scatterPlot(df = df_plot, var = "Ground_truth", colvec = col_vector[1:20]) 
plot_list[[3]] <- scatterPlot(df = df_plot, var = "asCIDER_cluster", colvec = col_vector[1:20]) 
plot_grid(plotlist = plot_list, ncol = 1)
```




# References

1. Baron, M. et al. A Single-Cell Transcriptomic Map of the Human and Mouse Pancreas Reveals Inter- and Intra-cell Population Structure. Cell Syst 3, 346–360.e4 (2016).
